### 2024-11-26 paper/al.tex
Minor fix
More consistent wording

% Execute
On the other hand, some control instructions \textit{execute} a Wasm
instruction.
\cref{fig:if} represents the prose specification for the \texttt{if}
instruction, which executes a block instruction with instructions in the then
branch or the else branch based on the value in the stack.
However, the block instruction itself does not appeared in the actual code,
making its execution unclear in this context.
The situation become even more complicated when the block contains control
instructions, such as function call.
While the intended meaning can still be inferred, this informality leaves room
for wrong interpretations.

\section{A Formal Model for WebAssembly Control Flow}
\label{sec:model}

% TODO: Clever way to distinguish jargon rather then \textit
% Formal model
To formally specify the prose specification, we construct a formal model the
prose specification.
% Execute
Rather than relying on a program counter, we employ an instruction stack to
represent execution progress.
Execution proceeds by popping an instruction from the stack and performing its
operation.
To avoid confusion with the stack that contains values interleaved with
contexts, we refer to the latter as the \textit{interleaved stack}.
This approach allows us to define the execute operation without introducing a
program counter.

% Jump
We then split two distinct usages of the jump operation: \textit{enter} and
\textit{exit}.
Since Wasm supports structured control flow, the jump operation is used either
for entering or exiting control structures such as a function body and a block
of instructions.
% Enter
We define the enter as the operation of pushing the context associated with the
jump onto the interleaved stack and the instructions within the structure onto
the instruction stack.
Similarly, we define exit as the operation of popping the corresponding context
and instructions.

% State transition
To express the enter and exit operation formally, we model them as state
transitions.
We define a Wasm state as a tuple of the interleaved stack, the instruction
stack, and the store.
For brevity, we omit the store in this discussion.
Let the interleaved stack initially contains $n$ entries $we_1 ~ ... ~ we_n$,
and the instruction stack contains $m$ entries $wi_1 ~ ... ~ wi_m$.
Suppose $wi_1$ is a control instruction that enters a control structure
containing a sequence of k instrucitons $wi_{11} ~ ... ~ wi_{1k}$ with an
associated context $wc$.
The enter operation can be expressed as the following state transition:
\[
(we_1 ~ ... ~ we_n, ~ wi_1 ~ ... ~ wi_m) \quad\leadsto\quad
  (
    wc ~ we_1 ~ ...  ~ we_n, ~
    wi_{11} ~ ... ~ wi_{1k} ~ \textbf{end}_{wc} ~ wi_2 ~ ... ~ wi_m
  )
\]

>>>>

% Execute
On the other hand, some control instructions \textit{execute} another Wasm
instruction.
\cref{fig:if} represents the prose specification for the \texttt{if}
instruction, which executes a block instruction with instructions in the then
branch or the else branch based on the value in the stack.
However, the block instruction itself does not appeared in the actual code,
making its execution unclear in this context.
The situation become even more complicated when the block contains control
instructions, such as function call.
While the intended meaning can still be inferred, this informality leaves room
for wrong interpretations.

\section{A Formal Model for the prose specification}
\label{sec:model}

% TODO: Clever way to distinguish jargon rather then \textit
% Formal model
To formally specify the prose specification, we construct a formal model for
the prose specification.
% Execute
Rather than relying on a program counter, we employ an instruction stack to
represent execution progress.
Execution proceeds by popping an instruction from the stack and performing its
operations.
To avoid confusion with the stack that contains values interleaved with
contexts, we refer to the latter as the \textit{interleaved stack}.
This approach allows us to define the execute operation without introducing a
program counter.
We define a Wasm state as a tuple of the interleaved stack, the instruction
stack, and the store.
An operation, as described in a step of the prose specification, is then
formalized as a state transition.

% Jump
To define the jump formally, we split two distinct usages of the jump
operation: \textit{enter} and \textit{exit}.
Since Wasm supports structured control flow, the jump operation is used either
for entering or exiting control structures such as a function body and a block
of instructions.
% Enter
We define the enter as the operation of pushing the context associated with the
jump onto the interleaved stack and the instructions within the structure onto
the instruction stack.
Similarly, we define exit as the operation of popping the corresponding context
and instructions.

% State transition
We then formalize the enter operation as a state transition.
For brevity, we omit the store in this discussion.
Let the interleaved stack initially contains $n$ entries $we_1 ~ ... ~ we_n$,
and the instruction stack contains $m$ entries $wi_1 ~ ... ~ wi_m$.
Suppose $wi_1$ is a control instruction that enters a control structure
containing a sequence of k instrucitons $wi_{11} ~ ... ~ wi_{1k}$ with an
associated context $wc$.
The enter operation can be expressed as the following state transition:
\[
(we_1 ~ ... ~ we_n, ~ wi_1 ~ ... ~ wi_m) \quad\leadsto\quad
  (
    wc ~ we_1 ~ ...  ~ we_n, ~
    wi_{11} ~ ... ~ wi_{1k} ~ \textbf{end}_{wc} ~ wi_2 ~ ... ~ wi_m
  )
\]
